---
title: "Gov 50 Final Project"
author: "Lydia Fraser"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to Chinatown-preferably Boston's Chinatown but am also open to doing research into one based in another city. I would also be interested in doing research on a suburban Asian enclave in the greater Los Angeles area of Southern California.

```{r}
knitr::opts_chunk$set(error = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(broom)
```

## Project Proposal
My research question is analyzing Asian voting behavior by looking at a specific suburban enclave in Los Angeles. I want to analyze the relationship between being Asian and voting based on descriptive representation versus partisan representation. Despite being a typically blue demographic, there have been several elections in which a Republican Asian candidate has won. My question is ultimately what the demographic shifts in the San Gabriel Valley from the last round of redistricting can reveal about Asian voting behavior through the most recent elections in 2022. My hypothesis is that Asians voters generally value descriptive representation over partisan representation when these identities come into conflict. I will do this by comparing census and election data from 2020 and 2022 from California District 38 (before and after redistricting and East San Gabriel Valley, a predominantly Asian area, was redistricted to CA-38). My variables of interest are votes cast for a given candidate in a particular election (either congressional or presidential). I will do multiple regressions to visualize and compare these votes. 

## Milestone 3
```{r}
votes <- read_csv("tract.csv")

cngdem22 <- votes |>
  filter(cd_2020 == "38")
  lm(CNGDEM01_22 ~ pop, data = cngdem22)
cngdem22

cngdem22_plot <- ggplot(cngdem22, mapping = aes(x = pop, y = CNGDEM01_22)) + geom_point() + geom_smooth(method = "lm") + scale_y_continuous(limits = c(0, 2500)) + labs(x = "Population", y = "Number of Votes", title = "2022 Congressional Votes for Linda Sánchez Regression")
cngdem22_plot
```

```{r}
cngdem1820 <- votes |>
  filter(cd_2010 == "38")
lm(CNGDEM01_18 ~ pop, data = cngdem1820)
lm(CNGDEM01_20 ~ pop, data = cngdem1820)
cngdem1820

cngdem1820_plot <- ggplot(cngdem1820, mapping = aes(x = pop)) + geom_point(aes(y = CNGDEM01_18, color = "2018")) + geom_smooth(aes(y = CNGDEM01_18), method = "lm", color = "blue") + geom_point(aes(y = CNGDEM01_20, color = "2020")) + geom_smooth(aes(y = CNGDEM01_20), method = "lm", color = "red")  + labs(x = "Population", y = "Number of Votes", title = "2018 vs. 2020 Congressional Votes for Linda Sánchez Regression") + scale_color_manual(name = "CNG Election", values = c("2018" = "blue", "2020" = "red"), labels = c("2018", "2020"))
cngdem1820_plot
```


The two plots are meant to demonstrate/confirm my preliminary observations about Asian voting behavior. The first graph visualizes a linear regression for votes for Linda Sánchez, the Democratic candidate, in 2022 (after East SGV was redistricted into CA-38). The second graph visualizes two linear regressions for votes for Sánchez in the 2018 and 2022 Congressional elections to compare the impact of redistricting and increase in Asian voters in CA-38. When juxtaposed, we can see that the difference in votes for Sánchez in 2022 versus 2018 and 2020, before the uptick in the number of Asians in CA-38. In 2022, with the presence of an Asian Republican candidate, the proportion of votes for Sánchez decreased, despite being an incumbent candidate. Because the other areas of CA-38 pre- and post-redistricting are kept relatively consistent, we can more comfortably attribute this shift in votes to the increase in Asian voters, who are usually a pretty consistent blue voting bloc.
